---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(ggplot2)
unzip("activity.zip")
activitydata <- read.csv("activity.csv", colClasses = c("numeric", "Date", "numeric"))

```

## What is mean total number of steps taken per day?

###Histogram of the sum of the steps based on date
```{r steps by date histogram}
sumsteps <- aggregate(steps~date, activitydata, sum, na.action = na.omit)
hist(sumsteps$steps, main = "Total Steps per Day", xlab = "Steps")
```

###Mean steps per date
```{r}
meansteps <- aggregate(steps~date, activitydata, mean)
print(meansteps)
```
###Median steps per date
```{r, echo=TRUE}
mediansteps <- aggregate(steps ~ date, activitydata, mean)
mediansteps
```

## What is the average daily activity pattern?
```{r}

meanstepsinterval <- aggregate(steps~interval, activitydata, mean)
```

###Average steps by interval across all days
```{r mean steps by interval}
ggplot(meanstepsinterval, aes(interval, steps))  + 
        geom_line() +
        ggtitle("Time Series of Mean Steps by Interval")
```

###Interval that saw the most average steps
```{r}

meanstepsinterval$interval[meanstepsinterval$steps == max(meanstepsinterval$steps)]
```

## Imputing missing values

###Number of rows with NA values
```{r}

sum(is.na(activitydata$steps))

```

###Replace NA step values with their mean at the interval
```{r}
repdata <- activitydata
for (i in 1:nrow(repdata)){
if (is.na(repdata$steps[i]) == T){
        #avgst <- meanstep
        repdata$steps[i] <- meanstepsinterval$steps[repdata$interval[i] == meanstepsinterval$interval]

        }
}
```

###Total number of steps by day with imputed data
```{r steps by date on replaced data}

hist(aggregate(steps~date, repdata, sum)$steps, main="Total Steps by Day", xlab = "Steps")

```
## Are there differences in activity patterns between weekdays and weekends?
###Create a new variable that determines if a day is a weekday or not
```{r}
for(i in 1:nrow(repdata)) {
if(weekdays(repdata$date[i])=="Saturday" | weekdays(repdata$date[i])=="Sunday"){
        repdata$daytype[i] <- "Weekend"
} else {
        repdata$daytype[i] <- "weekday"
}
}
  
```
###Average steps by interval by weekday or weekend
```{r average steps by interval by daytype}
aggrepdata <- aggregate(steps~interval+daytype, repdata, mean)
ggplot(aggrepdata, aes( interval, steps)) +
        geom_line() + facet_grid(daytype ~ . )
```
